---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
---

```{r libraries}
library(tidyverse)
library(forcats)
library(janitor)
```

```{r data}
funded <- read_csv("data/FundedSports.csv")
```
# Basic plot
This a basic blot, typical of ggplot's default values.
```{r}
basic_plot <- ggplot(data = funded) +
  geom_bar(aes(x = fct_reorder(Location, Year), 
               y = Funding,
               fill = fct_reorder(SportGroup, Funding)), 
           stat = "summary", fun.y = "sum")

basic_plot
```
# Setting labels
There is a function in ggplot, labs. Can set

* title, subtitle, caption

* x, y

* colour, fill, shape

```{r setting-labels}
label_plot <- basic_plot +
  labs(title = "Changes in Funding For Olympic Sports",
       subtitle = "Funding Provided by UK Sport World Class Performance Programme",
       x = "", y = "", 
       caption = "Data taken from uksport.gov.uk",
       fill = "Event")

label_plot
```

```{r}
ex_plot <- ggplot(data = funded) +
  geom_line(aes(x = Date,
                y = Funding,
                colour = Sport))
            
ex_plot_labs <- ex_plot +
  labs(title = "Funding over time for Summer Olympic sports",
       subtitle = "As provided by UK Sport",
       x = "", y = "")

ex_plot_labs

```
# Scales

Naming convention:
```

```
Control:

* Break points (tick marks)

* Labels

* Limits

* Colour/shape/size

* ...

Scale functions let us define how the X and Y axes appear

We can use scale functions for:

* Continuous

* Discrete

* Date/Time

* Transformation
```{r}
labels <- c("Sydney\n(2000)", "Athens\n(2004)", 
            "Beijing\n(2008)", "London\n(2012)", 
            "Rio de Janeiro\n(2016)")

axis_plot <- label_plot +
  scale_x_discrete(labels = labels) +
  scale_y_continuous(breaks = seq(0, 25e7, by = 50e6),
                     minor_breaks = NULL,
                     labels = scales::dollar_format("£"))

axis_plot

```

## Colours, shapes and sizes
The default options are:

* Continuous for numeric data

* Manual for setting specific values

There are a number of predefined options (particularly for colour)
* gradient

* brewer

* ...

```{r}
cols <- c( "grey", "lightBlue", "orangeRed", 
           "steelBlue", "royalBlue", "navyBlue")

axis_plot_cols <- axis_plot +
  scale_fill_manual(values = cols)

axis_plot_cols
```

```{r}
cities <- c("Sydney", "Athens", 
            "Beijing", "London", 
            "Rio de Janeiro")

ex_plot_months <- ex_plot_labs +
  scale_x_date(breaks = unique(funded$Date),
                 date_labels = paste0(cities, "\n%b %Y")) +
  scale_y_continuous(labels = scales::dollar_format("£"))

ex_plot_months
```

# Legends
## Themes
Position of the legend is controlled by the themes layer
## Guides
The guide functions let us control all aspects of a legend

* titles

* positioning

* ordering
```{r}
axis_plot_legend <- axis_plot_cols +
  theme(legend.position = "bottom",
        legend.justification = "left") +
  guides(fill = guide_legend(
    title.position = "top",
    title.hjust = 0.5,
    nrow = 1,
    label.position = "bottom"
  ))

axis_plot_legend
```

```{r}
ex_plot_legend <- ex_plot_months +
  theme(legend.position = "bottom",
        legend.justification = "left") +
  guides(colour =
           guide_legend(
            override.aes = list(size = 4,
                                alpha = 0.5),
            nrow = 4,
            title.position = "top"
    )
  )

ex_plot_legend
```

```{r}
cont_plot <- ggplot(data = funded) +
  geom_bar(aes(x = fct_reorder(Location, Year),
               y = Funding, fill = Year),
           stat = "summary", fun.y = "sum") +
  guides(fill = guide_legend())

cont_plot
```

# Themes
There are a number of default themes that will control the look and feel of our plot

We can add themes for a session or on a plot by plot basis

* After adding for a session

```{r}
theme_set(theme_classic(base_size = 14))
```

# Custom themes
* We can create a custom theme using the `theme_update` function

```{r}
theme_update(legend.position = "bottom")

cont_plot

```

# Element types
All aspects of a graphic are either
* rectangle

* line

* text

We set them with the appropriate `element_*` function

We can configure:
* colour/fill

* size

* justification & angle

* fonts

```{r}
theme_update(legend.position = "bottom",
             axis.text.y = element_text(angle = 45))

axis_plot_cols

```

```{r}
theme_update(legend.position = "bottom",
             axis.text.y = element_text(angle = 45),
             panel.grid.major.y = element_line(colour = "black",
                                               #dashed
                                               linetype = 2),
             panel.background = element_rect(fill = "turquoise"))

axis_plot_cols

# axis_plot_cols +
  # theme()

```

```{r}
theme_set(theme_bw())

theme_update(panel.grid.major.y = element_line(linetype = 2),
             plot.title = element_text(hjust = 0.5),
             plot.subtitle = element_text(hjust = 0.5))


ex_plot_legend

```

```{r}
theme_update(legend.position = "bottom")

cont_plot +
  theme(axis.title.x = element_text(margin = margin(40, 0, 0, 0)))

```